# 3. Sports Cards

In this experiment, the experimenters invited consumers at a sports card trading show to bid against one other bidder for a pair trading cards.  We abstract from the multi-unit-auction details here, and simply state that the treatment auction format was theoretically predicted to produce lower bids than the control auction format.  We provide you a relevant subset of data from the experiment.

In this question, we are asking you to produce p-values and confidence intervals in three different ways: 

1. Using a `t.test`; 
2. Using a regression; and,
3. Using randomization inference. 

```{r load cards data }
d <- fread('../data/list_data_2019.csv')
```

1. Using a `t.test`, compute a 95% confidence interval for the difference between the treatment mean and the control mean. After you conduct your test, write a narrative statement, using inline code evaluation that describes what your tests find, and how you interpret these results. (You should be able to look into `str(t_test_cards)` to find the pieces that you want to pull to include in your written results.) 

```{r cards t-test}
t_test_cards <- d[ , t.test(bid ~ uniform_price_auction)]
t_test_cards
```

2. In plain language, what does this confidence interval mean? 

> The confidence interval that is reported is `r as.numeric(round(t_test_cards$conf.int, 2))`. This means that if the experiment were re-conducted with the same parameters and population, this range would cover the true value in 95 / 100 of the re-conducted experiments. 

3. Conduct a randomization inference process using an estimator that you write by hand (i.e. in the same way as earlier questions). On the sharp-null distribution that this process creates, compute the 2.5% quantile and the 97.5% quantile using the function `quantile` with the appropriate vector passed to the `probs` argument. After you conduct your test, write a narrative statement of your test results. 

```{r cards randomization inference} 
## first, do you work for the randomization inference
n_ri_loops <- 1000

cards_ate <-  d[ , 
  .(mean_bid = mean(bid)), 
  keyby = (uniform_price_auction)][ , 
    diff(mean_bid)]


cards_ri_distribution <- replicate(
  n = n_ri_loops, 
  expr = d[ ,             
    .(mean_bid = mean(bid)), 
    keyby = (sample(uniform_price_auction))][ ,
      diff(mean_bid)]
  )


cards_ri_quantiles <- quantile(cards_ri_distribution, c(0.025, 0.975))
cards_ri_quantiles

# I'm choosing to use two sided p-values. Because I'm not a monster. 
cards_ri_p_value <- mean(abs(cards_ri_distribution >= cards_ate))
cards_ri_p_value
```

> The thing is, producing a confidence interval doesn't really provide a test, right? To conduct that test, we have to go through the process of comparing the ri distribution to the estimated ate. 
> 
> However, you should be able to see that this permutation based method is generating something similar to the t-test that you computed. 

```{r}
cards_ate + cards_ri_quantiles
```

> This isn't a tightly principled way to reason. In RI, there is no uncertainty about the estimated treatment effect -- it simply is what it is. And, the uncertainty is placed around the values that could have arisen if the null hypothesis were actually true.

4. Do you learn anything different if you regress the outcome on a binary treatment variable? To answer this question, regress `bid` on a binary variable equal to 0 for the control auction and 1 for the treatment auction and then calculate the 95% confidence interval using *classical standard errors* (in a moment you will calculate with *robust standard errors*). There are two ways to do this -- you can code them by hand; or use a built-in, `confint`. After you conduct your test, write a narrative statement of your test results. 

```{r cards ols regression}
mod <- d[ , lm(bid ~ uniform_price_auction)]
```

5. Calculate the 95% confidence interval using robust standard errors, using the `sandwich` package. There is a function in `lmtest` called `coefci` that can help with this. It is also possible to do this work by hand. After you conduct your test, write a narrative statement of your test results.

```{r cards robust ci}
cards_robust_ci <- coefci(mod, vcov. = vcovHC(mod))
```

> This robust confidence interval is _remarkably_ similar to the CI that I generated with the t-test; and, if I squint a little bit at the randomization-inference, it also seems to be covering a pretty similar range. Although the interpretation differs between the RI and frequentist (i.e. t-test and OLS) estimates, the upshot of each of these seems to be quite simialar. 

6. Characterize what you learn from each of these different methods -- are the results contingent on the method of analysis that you choose? 

> What we hope you take away from this is that _when the data meets the assumptions of all the methods, they should produce **very** similar estimates and tested values._ It is when the data presents a challenge to one of the methods that we have to elect to use a method that is consistent with the requirements of that method. 
